Fig2: Candidate genes

### 2020-04-17 analysis/figure-2.Rmd
now with candidate genes as fig 2

```{r}
library(tidyverse)
library(ggtext)
library(cowplot)
library(ggpubr)
library(knitr)
library(kableExtra)

@@

table1 <- left_join(candidateDEGS, GOgenesWide) %>%
  select(gene, bldg_lay:n5_n9, parentalcare, parentalbehavior, NCBI, ) %>%
  mutate(parentalcare = if_else(is.na(parentalcare), " ", "X"),
         parentalbehavior = if_else(is.na(parentalbehavior), " ", "X")) %>%
  rename("Literature" = "parentalcare", "GO" =  "parentalbehavior")
table1[is.na(table1)] <- " "
table1

@@

```{r eval = F}

@@

head(allvsd)
candidategenes <- GOgenesLong %>% distinct(gene) %>% pull(gene)

@@

```{r fig2, fig.width = 7, eval = F}

newboxplot <- function(df, mygenes, whichsex, whichstage){

  p <- df %>%
    filter(gene %in% mygenes, 
           sex %in% whichsex,
           treatment %in% whichstage) %>%
    droplevels() %>%
    mutate(tissue = factor(tissue, levels = tissuelevel)) %>%
    ggplot(aes(x = treatment, y = counts, fill = treatment, color = sex)) +
    geom_boxplot() +
    geom_jitter(size = 0.5, width = 0.1) +
    facet_grid(tissue~gene, scales = "free_y") +
    theme_B3() +
    scale_color_manual(values = allcolors) +
    scale_fill_manual(values = allcolors) +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 45, hjust = 1),
          strip.text.x = element_text(face = "italic"),
          axis.title.x = element_blank())  +
    labs(x = "Parental stage",
         y = "variance stabilized expression")
  return(p)
}

a1 <- newboxplot(hypvsd, c("PGR",  "CYP19A1", "LHCGR" ), "female", c( "hatch", "n5")) + 
  theme(strip.text.y = element_blank()) + labs(subtitle = "Females")
a2 <- newboxplot(hypvsd, c("AR"), "male", c( "hatch", "n5")) + 
  theme(axis.title.y = element_blank()) + labs(subtitle = "Males")

b1 <- newboxplot(gonvsd, c("HSD3B2"), "female", c( "bldg", "lay"))  +
  theme(strip.text.y = element_blank(),
        axis.title.y = element_blank()) + labs(subtitle = "Females")
b2 <- newboxplot(gonvsd, c("VIPR1",  "PRLR" ), "female", c( "lay", "inc.d3"))  + 
  theme(axis.title.y = element_blank()) + labs(subtitle = " ")

ab <- plot_grid(a1,a2,b1,b2, rel_widths = c(6,2.5,2,4), labels = c("a", " ", "b", " "), label_size = 8, nrow = 1)

c1 <- newboxplot(pitvsd, c( "ESR1", "GNRHR"), "female", c("bldg", "lay",  "inc.d3")) 
c2 <- newboxplot(pitvsd, c( "PRL"), "female", c("inc.d9", "inc.d17", "hatch", "n5")) 
c3 <- newboxplot(pitvsd, c("VIPR1",  "PRL" ), "male", c( "inc.d9", "inc.d17"))

c <- plot_grid(c1 + theme(strip.text.y = element_blank()) + labs(subtitle = "Females"),
          c2 + theme(strip.text.y = element_blank(),
                    axis.title.y = element_blank()) + labs(subtitle = " ") ,
          c3 + theme(axis.title.y = element_blank()) + labs(subtitle = "Males"), 
          nrow = 1, rel_widths = c(6,4,4),
          labels = c("c"), label_size = 8)

plot_grid(ab,c, nrow = 2)

# correlations
library(corrr)

hormones <- read_csv("../results/07_hormoneswide.csv") %>%
  select(-study, -treatment, -sex)
head(hormones)

head(candidatevsd)

vsdhormones <- candidatevsd %>%
  pivot_wider(names_from = gene, values_from = counts) %>%
  mutate(bird_id = sapply(strsplit(samples,"\\_"), "[", 1)) %>%
  left_join(.,hormones, by =  "bird_id")
vsdhormones

vsdhormoneshyp <- vsdhormones %>% filter(tissue == "hypothalamus" )
vsdhormonespit <- vsdhormones %>% filter(tissue == "pituitary" )
vsdhormonesgon <- vsdhormones %>% filter(tissue == "gonad" )

plotcorrs <- function(whichtissue, whichsex){

  x <- vsdhormones  %>%
    filter(tissue == whichtissue, 
           sex == whichsex) %>%
    select(-sex, - tissue, - treatment, -samples, -bird_id, -moltbin) %>%
    correlate()
  print(fashion(x))

  p <-rplot(x, colors = c("#0571b0", "#92c5de", "#f7f7f7","#f7f7f7","#f7f7f7", "#f4a582", "#ca0020")) +
    theme_B3() +
    theme(legend.position = "bottom",
          axis.text.x = element_text(angle = 45, hjust = 1),
          axis.text = element_text(face = "italic")) +
    labs(subtitle = whichtissue)
  return(p)
}

a <- plotcorrs("hypothalamus", "female") + theme(legend.position = "none")
b <- plotcorrs("pituitary", "female") + theme(axis.text.y = element_blank()) + theme(legend.position = "none")
c <- plotcorrs("gonad", "female") + theme(axis.text.y = element_blank()) + theme(legend.position = "none")

d <- plotcorrs("hypothalamus", "male") 
e <- plotcorrs("pituitary", "male") + theme(axis.text.y = element_blank())
f <- plotcorrs("gonad", "male") + theme(axis.text.y = element_blank())

plot_grid(a,b,c, d,e,f, nrow = 2, rel_widths = c(1.2,1,1), rel_heights = c(1,1.2),
          labels = c("females", " ", " ",
                     "males", " ", ""),
          label_size = 8)

plotspecificcorrs <- function(df, myx, myy, myxlab, myylab){
  p <- df %>%
    mutate(tissue = factor(tissue, levels = tissuelevel)) %>%
    ggplot(aes(x = myx, y = myy)) +
    geom_point(aes(color = treatment)) +
    geom_smooth(method = "lm", aes(color = sex)) +
    facet_wrap(~sex, scales = "free", nrow = 2) +
    scale_color_manual(values = allcolors) +
    labs(x = myxlab, y = myylab) +
    theme_B3() +
    theme(legend.position = "none",
          axis.title = element_text(face = "italic"))
  return(p)
}

a <- plotspecificcorrs(vsdhormoneshyp, vsdhormoneshyp$FSHB, vsdhormoneshyp$PRL, "FSHB", "PRL") + labs(subtitle = "Hypothalamus")
b <- plotspecificcorrs(vsdhormonespit, vsdhormonespit$prolactin, vsdhormonespit$PRL, "circulating prolactin", "PRL") +
  theme(axis.title.x = element_text(face = "plain")) + labs(subtitle = "Pituitary")
c <- plotspecificcorrs(vsdhormonesgon, vsdhormonesgon$ESR1, vsdhormonesgon$PGR, "ESR1", "PGR")  + labs(subtitle = "Gonads")

plot_grid(a,b,c,nrow = 1, labels = "auto", label_size = 8)
```

>>>>

```{r setup}
library(tidyverse)
library(ggtext)
library(cowplot)
library(ggpubr)
library(knitr)
library(kableExtra)

@@

table1 <- left_join(candidateDEGS, GOgenesWide) %>%
  select(gene, bldg_lay:n5_n9, parentalcare, parentalbehavior, NCBI) %>%
  mutate(parentalcare = if_else(is.na(parentalcare), " ", "X"),
         parentalbehavior = if_else(is.na(parentalbehavior), " ", "X")) %>%
  rename("Literature" = "parentalcare", "GO" =  "parentalbehavior")
table1[is.na(table1)] <- " "
table1

@@

```{r vsds}

@@

@@

```{r fig2, fig.width = 7.25, fig.height = 7.25}

# significant candidate genes
hypDEGs <- c("CRHBP", "CRHR2", "DRD1", "GNAQ", 
             "HTR2C", "OPRK1", "PGR")
pitDEGs <- c("ESR1", "MEST", "PRL", "PTEN")
gonadDEGs <- c("AVPR1A", "CREBRF", "FOS", "GNAQ", 
               "NR3C1", "OPRM1", "PRLR")

makelineplots <- function(whichtissue, whichgenes){

  p <- candidatevsd %>%
    filter(tissue == whichtissue,
          gene %in% whichgenes) %>%
    mutate(treatment = factor(treatment, levels = charlevels)) %>%
    mutate(treatmentNum = as.numeric(treatment)) %>%
    ggplot(aes(x = treatmentNum, y = counts)) +
    geom_boxplot(aes(fill = treatment, color = sex), outlier.shape = NA) +
    #geom_smooth(aes(color = sex)) +
    geom_jitter(size = 0.5, aes(color = sex)) +
    facet_grid(gene~sex, scales = "free") +
    scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8,9),
                       labels = charlevels) +
    scale_fill_manual(values = allcolors) +
    scale_color_manual(values = allcolors) +
    theme_B3() + 
    theme(legend.position = "none",
          strip.text.y = element_text(face = "italic"),
          axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(subtitle = whichtissue, 
         y = "gene expression (variance stabilized)",
         x = "Parental Care Stage") 
  return(p)
}

a <- makelineplots("hypothalamus", hypDEGs)
b <- makelineplots("pituitary", pitDEGs) + labs(y = NULL)
c <- makelineplots("gonad", gonadDEGs) + labs(y = NULL)

plot_grid(a,b,c, nrow = 1, labels = "AUTO", label_size = 8)
```
